---
title: "Introduction to opcr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to opcr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = 'center',
  fig.width = 6, 
  fig.height = 4
)
```

```{r setup}
library(opcr)
```

## Process

### Reading raw data

Reading and processing the data is all done using `opc_process()`. This does the following steps:  

1. read in binary OPC data in .D00 file with `read_focal_opc()`  
2. convert to a nice tabular format with `convert_single_opc()`  
3. apply various quality control flags with `opc_flag()`  
4. use a shiny app to interactively select the downcast with `opc_trim()`  

Here's a simple example of how to process the raw OPC data. Note that this requires the user to interactively select the downcast so does not actually run in this vignette.
```{r process, eval = FALSE}
opc = opc_process(ifile = 'data/raw/OPC027.D00')
```

### OPC data format

The processed data is stored in what I refer to as an `opc tibble` with the following fields:  

`scan` - the scan number, which iteratively increases with each data record  
`timer` - the timer, which recorded seconds since the unit was powered on  
`atten` - light attenuation  
`depth` - instrument depth in meters  
`flag` - quality control flag, with zero meaning good. See `opc_flag()` for the other definitions  
`time` - the datetime since deployment, in UTC  
`volume_filtered` - the volume of water that has passed through the OPC since the previous record, in cubic meters  
`esd` - a nested list of the particle sizes, in Equivalent Spherical Diameter (ESD; mm) detected during this data record  

Here's an example:  
```{r}
opc
```

Data in this format is the cornerstone of this package; all computation and plotting requires this `opc tibble` as input.

## Calculate

Several functions make it convenient to calculate common OPC data products.

### Size-frequency histogram

```{r}
opc_histogram(opc)
```

### Abundance profile

```{r}
opc_abundance(opc)
```

### Biomass profile

This uses the equation from Suthers et al 2006 to covert size to biomass.

```{r}
opc_biomass(opc)
```

### Abundance in size and depth matrix

```{r}
opc_image(opc)
```

## Plot

Here are some quick plots. These take an `opc tibble` as input, derive the relevant variables using the functions above, then return a nice ggplot. Check the function documentation for more information.

### Simple downcast

```{r}
opc_plot_depth(opc)
```

### Flags

```{r}
opc_plot_flags(opc)
```

### Attenuation

```{r}
opc_plot_attenuance(opc)
```

### Histogram

```{r}
opc_plot_histogram(opc)
```

### Abundance

```{r}
opc_plot_abundance(opc)
```

### Biomass

```{r}
opc_plot_biomass(opc)
```

### Image

```{r, fig.width=6,fig.height=6}
opc_plot_image(opc)
```

### Multipanel

```{r,fig.width=6,fig.height=6}
opc_plot_multipanel(opc)
```

### Diagnostics

```{r,fig.width=8,fig.height=6}
opc_plot_diagnostics(opc)
```
